---
title: "KHN Data"
output: html_notebook
---

Run packages
```{r}
#install.packages("tidyverse")
library("tidyverse")
#install.packages("rio")
library("rio")
#install.packages("pacman")
library("pacman")
#install.packages("rmarkdown")
library("rmarkdown")
```

Upload data and create a dataframe called khn
```{r}
download.file("https://raw.githubusercontent.com/khnews/2020-underfunded-under-threat-data/master/data/05-local-health-departments-detail.csv", "05-local-health-departments-detail.csv")
khn <- rio::import("data/05-local-health-departments-detail.csv")

```

Create a new df called nc that filters just for nc
```{r}
nc <- khn %>%
  filter(state_code == "NC")
```

It looks like some single counties may have been incorporated into a multi county system after '99. This was the case with Polk. I imagine that this is the case for counties that show "1" in the total count
```{r}
single_counties <- nc %>%  
  filter(lhd_area_type == "single county") %>%  
 group_by(lhd_name) %>%
  summarise(
    Total = n()
  )

filter(single_counties, Total == "1")
```

What are the multi counties?
```{r}
multi_counties <- nc %>%  
  filter(lhd_area_type == "multi-county") %>%  
 group_by(lhd_name) %>%
  summarise(
    Total = n()
  )
```
I did this code below without accounting for the multi-county systems

What is the average number of full time employees for each county in NC from 1999-2017? 
```{r}
nc_fte <- nc %>%
  filter(state_code == "NC", !is.na(fte)) %>%
  group_by(lhd_name) %>%
  summarise(
    AverageFte = mean(fte)
    ) %>%
  select(lhd_name, AverageFte) %>%
  arrange(AverageFte)
```
What was the average fte for each county in 2017? 
```{r}
nc_fte_2017 <- nc %>%
  filter(year == "2017", !is.na(fte)) %>%
  group_by(lhd_name) %>%
  summarise(
    AverageFte = mean(fte)
    ) %>%
  select(lhd_name, AverageFte) %>%
  arrange(AverageFte)
```
The average of those averages above? How does it compare to the highest fte from the previous code? Lowest?
```{r}
nc_fte_2017 %>%
  mutate(
    Avg = mean(AverageFte)
  ) %>%
  select(Avg)

nc %>%
  filter(year == "2017", !is.na(fte)) %>%
  group_by(lhd_name) %>%
  summarise(
    AverageFte = mean(fte)
    ) %>%
  select(lhd_name, AverageFte) %>%
  arrange(desc(AverageFte))

#97.8 vs 825 (Mecklenburg)

nc %>%
  filter(year == "2017", !is.na(fte)) %>%
  group_by(lhd_name) %>%
  summarise(
    AverageFte = mean(fte)
    ) %>%
  select(lhd_name, AverageFte) %>%
  arrange(AverageFte)

#97.8 vs 14 (Hyde)
```


